---
title: "SVM Model and Naive Bayes"
output:
  html_document: default
  html_notebook: default
---

## Problem Statement
The data we will be using in this analysis is data from Twitter on airlines. I will do a sentiment analysis on tweets and classify whether they are postive, negative or neutral.

I will make use of the bag of words representation provided by the TextMining package tm and will train with SVM and a multinomial Naive Bayes classifier using and the e1071 package.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Needed <- c("tm", "SnowballCC")
# install.packages(Needed, dependencies=TRUE)
# install.packages("Rcampdf", repos = "http://datacube.wu.ac.at/", type = "source")
library(tm)
library(e1071)
library(SnowballC)
```

## Data Preparation

Import and shuffle the data
```{r, message=FALSE, warning=FALSE}
setwd('/Users/charlotteleysen/Google Drive/*PROJECTS/IE/Term 2/Machine Leaning 2/SVM_Naive Assignment/assignment_data/charlotte_leysen_assignment/')
df <- read.csv('training.csv')
df_test <- read.csv('test.csv')
# Randomize the dataset to facilitate the training process
set.seed(123)
df <- df[sample(nrow(df)), ]

# Convert the target variable ('class') from character to factor.
df$airline_sentiment <- as.factor(df$airline_sentiment)
```

Convert the data to a SimpleCorpus format
```{r, echo=TRUE}
corpus <- Corpus(VectorSource(df$text))

# We can take a look to the corpus
inspect(corpus[1:2])
```

Now I will create a function to clean up the text. The methods decided to apply are:

  * stemming
  * remove punctuation
  * strip blanks
  * remove stopwords

```{r}
cleanCorpus <- function(corpus) {
  corpus <-tm_map(corpus, stemDocument)
  corpus.tmp <- tm_map(corpus,removePunctuation)
  corpus.tmp <- tm_map(corpus.tmp,stripWhitespace)
  corpus.tmp <- tm_map(corpus.tmp,removeWords,stopwords("en"))
  return(corpus.tmp)
}
```

Using this function, I clean the corpus
```{r}
corpus.clean <- cleanCorpus(corpus)
```

Represent the bag of words tokens with a document term matrix (DTM). The rows of the DTM will correspond to the documents in the collection, columns to the terms, and its elements are the term frequencies.
```{r}
dtm <- DocumentTermMatrix(corpus.clean)

```


### Training and Test data sets.

Separate data into training and test sets (80% for training and 20% for test).

```{r}
dataset.train <- df[1:1500,]
dataset.test <- df[1501:2000,]

dtm.train <- dtm[1:1500,]
dtm.test <- dtm[1501:2000,]

corpus.clean.train <- corpus.clean[1:1500]
corpus.clean.test <- corpus.clean[1501:2000]
```

Now we are ready to start modelling the data.

# SVM Modelling

Import the caret library
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(caret)
```

Prepare the train and test data for SVM
```{r}

svmX <- as.matrix(dtm.train)
svmy <- dataset.train$airline_sentiment

training_data <- as.data.frame(cbind(svmy,svmX))
test_data <- as.data.frame(as.matrix(dtm.test))
```

Run the model

Using the functions in `e1071` package to create an SVM model for the training data, we train the model
```{r warning=FALSE}
sv <- svm(svmy~., training_data, type="C-classification", kernel="sigmoid", cost=1)
```

I evaluate the SVM model in terms of accuracy by predicting on the test data and computing the confusion matrix
```{r}
prediction_sv <- predict(sv, test_data)
table("Predictions"= prediction_sv,  "Actual" = dataset.test$airline_sentiment )

```

Calculate the accuracy from the confusion matrix
```{r}
acc_sv <- function(table){
  TP = table[1,1];  # true positives
  TN = table[2,2];  # true negatives
  FP = table[1,2];  # false positives
  FN = table[2,1];  # false negatives
  acc = (TP + TN)/(TP + TN + FP + FN)
  return(acc)
}
acc_sv(table("Predictions"= prediction_sv,  "Actual" = dataset.test$airline_sentiment ))
```
The result is 0.767 accuracy.

Now we try to tune SVM parameters to further improve the model performance using 5 fold cross validation. This code takes a while to run so have commented it out.
```{r warning=FALSE}
# fitControl <- trainControl(method = "cv",
#                            number = 5,
#                            verboseIter = TRUE)
# 
# cv.svm <- train(svmX,svmy,
#                 method="svmRadial",
#                 preProc = c("center", "scale"),
#                 tuneLength = 5,
#                 metric = "Accuracy",
#                 trControl = fitControl)
# 
# cv.svm.prediction <- predict(cv.svm, test_data)
# table("Predictions"= cv.svm.prediction,  "Actual" = dataset.test$airline_sentiment )
# 
# acc_sv(table("Predictions"= cv.svm.prediction,  "Actual" = dataset.test$airline_sentiment ))
```
The result is a small improvement on the original accuracy to 0.7899761.

Now I will model the data using Naive Bayes method to see whether we can find a better model.

# Naive Bayes Modelling 

### Feature Selection
The term-frequency matrix contains almost 10,000 features (i.e., terms). Not all of them are useful for classification so we will remove some.

```{r}
dim(dtm.train)
```

The removeSparseTerms function in the tm package helps to remove the sparse terms in the matrix.

```{r}
dtm.train.nb <- removeSparseTerms(dtm.train, 0.99)
```

I apply the same procedure for test set
```{r}
dtm.test.nb <- removeSparseTerms(dtm.test, 0.99)
```

For sentiment classification, word occurrence matters more than word frequency. Therefore, I replace term frequencies in the dataset by Boolean presence/absence features.
```{r}
# Convert the word frequency to binary 1/0 presence
binarize <- function(dataset) {
  factor(ifelse(dataset > 0, 1,0), levels=c(0,1), labels=c("Yes", "No"))
}

# Apply the function to training and test data
dtm.train.nb.binary <- apply(dtm.train.nb, 2, binarize)
dtm.test.nb.binary <- apply(dtm.test.nb, 2, binarize)
```

### Naïve Bayes Model
Making use of the `naiveBayes` package, I predict the polarity of the reviews in the test set.

```{r}
nb_model <- naiveBayes(dtm.train.nb.binary, dataset.train$airline_sentiment, laplace = 1)
probs <- predict(nb_model, newdata=dtm.test.nb.binary, type = "raw")
classes <- predict(nb_model, newdata=dtm.test.nb.binary, type = "class")
```

### Model evaluation

#### Confusion Matrix

I calculate the confusion matrix and the accuracy of the model. 
```{r}
# Confusion matrix
table("Predictions"= classes,  "Actual" = dataset.test$airline_sentiment )
```

Using the confusion matrix, I compute the Accuracy

```{r}
acc <- function(table){
  TP = table[1,1];  # true positives
  TN = table[2,2];  # true negatives
  FP = table[1,2];  # false positives
  FN = table[2,1];  # false negatives
  acc = (TP + TN)/(TP + TN + FP + FN)
  return(acc)
}
acc(table("Predictions"= classes,  "Actual" = dataset.test$airline_sentiment ))
```
The accuracy with Naive Bayes is 0.78. This is around the same as that of SVM with 5 fold cv. Since Naive Bayes is much faster, I will use this model for the final test data.

#Train the full data

Now I have chosen Naive Bayes for our model, I will retrain the model with all the data.
``` {r}
dtm.nb.all <- removeSparseTerms(dtm, 0.99)
dtm.nb.binary.all <- apply(dtm.nb.all, 2, binarize)
nb_model_all <- naiveBayes(dtm.nb.binary.all, df$airline_sentiment, laplace = 1)

```

### Clean and prepare the test data
``` {r}
corpus_test <- Corpus(VectorSource(df_test$text))
corpus.clean.test <- cleanCorpus(corpus_test)
dtm_test <- DocumentTermMatrix(corpus.clean.test)
dtm.nb.test <- removeSparseTerms(dtm_test, 0.99)
dtm.nb.binary.test <- apply(dtm.nb.test, 2, binarize)
```

### Predict the test data
```{r}
classes_test <- predict(nb_model_all, newdata=dtm.nb.binary.test, type = "class")
submission <- data.frame(cbind(df_test$tweet_id, classes_test))
```

### Prepare for submission
```{r}
submission[submission$classes_test == 1,"classes_test"] <- "negative"
submission[submission$classes_test == 2,"classes_test"] <- "neutral"
submission[submission$classes_test == 3,"classes_test"] <- "positive"
colnames(submission) <- c("tweet_id","airline_sentiment")

head(submission)

write.csv(submission, "/Users/charlotteleysen/Google Drive/*PROJECTS/IE/Term 2/Machine Leaning 2/SVM_Naive Assignment/assignment_data/charlotte_leysen_assignment/submission_file.csv", row.names = FALSE)

```



